---
title: "Functions in SamplingStrata package - III"
author: "Marco Ballin, Giulio Barcaroli"
date: "20 May 2019"
output: slidy_presentation
footer: "SamplingStrata Tutorial (uRos2019 - Bucharest 20-21 May 2019)"
incremental: true
widescreen: true
smaller: true
---


```{r, include = F}
library(SamplingStrata)
data("swissmunicipalities")
options(width=120)
```

## Use of models to take into account the anticipated variance

It has been assumed that, when optimizing the stratification of a sampling frame,
values of the target variables Y’s are available for the generality of the units in the frame, or at least for a
sample of them by means of which it is possible to estimate means and standard deviation of Y’s in atomic
strata. 

Of course, this assumption is seldom expected to hold. The situation in which some proxy variables are
available in the frame is much more likely to happen. 

In these situations, instead of directly indicating the real target variables, proxy ones are named as Y’s. 

By so doing, there is no guarantee that the final stratification
and allocation can ensure the compliance to the set of precision constraints.

In order to take into account this problem, and to limit the risk of overestimating the expected precision levels
of the optimized solution, it is possible to carry out the optimization by considering, instead of the expected
coefficients of variation related to proxy variables, the anticipated coefficients of variation (ACV) that depend on
the model that is possile to fit on couples of real target variables and proxy ones. In the current
implementation, only models linking continuous variables can be considered.

## Use of models to take into account the anticipated variance

In order to introduce the anticipated variance in the overall optimization, it is necessary first of all to model the target variable (dependent variable) with respect to the available proxy variable (independent variable).

For instance, let us consider the swiss municipalities example, 
and let have the same stratification variables and same target variables:

- "Surfacesbois","Surfacescult"
- "Pop020", "Pop2040"

But suppose now that the target variables are not available. Instead, only the area covered by buildings in each municipality is available ("Airbat"). 

Let us suppose to model the relationships between these variables for instance having the availability of a subset of observations:

```{r, eval = T}
set.seed(123)
mun_samp <- swissmunicipalities[sample(1:nrow(swissmunicipalities),500),]
```


## Use of models to take into account the anticipated variance

Now we can fit the model between "Pop020" and "Airbat" in this way:


```{r, eval = T}
mod_Pop_020_Airbat <- lm(log(Pop020) ~ log(Airbat), data=mun_samp)
summary(mod_Pop_020_Airbat)
```

## Use of models to take into account the anticipated variance

and between "Pop2040" and "Airbat" in this way:

```{r, eval = T}
mod_Pop_2040_Airbat <- lm(log(Pop2040) ~ log(Airbat), data=mun_samp)
summary(mod_Pop_2040_Airbat)
```

## Use of models to take into account the anticipated variance

We can now define the following dataframe:

```{r, eval = T}
model <- NULL
model$beta[1] <- mod_Pop_020_Airbat$coefficients[2]
model$sig2[1] <- summary(mod_Pop_020_Airbat)$sigma
model$type[1] <- "loglinear"
model$gamma[1] <- 0
model$beta[2] <- mod_Pop_2040_Airbat$coefficients[2]
model$sig2[2] <- summary(mod_Pop_2040_Airbat)$sigma
model$type[2] <- "loglinear"
model$gamma[2] <- 0
model <- as.data.frame(model)
model
```

## Optimization without taking into account the models

We define the frame in this way:


```{r, eval = T}
swissmunicipalities$id <- c(1:nrow(swissmunicipalities))
swissmunicipalities$dom <- 1

swissframe <- buildFrameDF(df = swissmunicipalities,
                           id = "id",
                           X = c("Surfacesbois",
                                 "Surfacescult"),
                           Y = c("Airbat",
                                 "Airbat"),
                           domainvalue = "dom")
swissframe$X1 <- var.bin(swissframe$X1, bins=15)
swissframe$X2 <- var.bin(swissframe$X2, bins=15)
```


and the precision constraints


```{r, eval = T}
swisserrors <- as.data.frame(list(DOM="DOM1",
                                  CV1=0.08,
                                  CV2=0.12,
                                  domainvalue=1
                            ))

swisserrors
```

## Optimization without taking into account the models

The optimization step:


```{r, eval = T}
swissstrata1 <- buildStrataDF(swissframe, progress = FALSE)

set.seed(123)
solution1 <- optimizeStrata(
  errors = swisserrors, 
  strata = swissstrata1,
  iter = 50,
  pops = 10,
  writeFiles = FALSE,
  showPlot = FALSE,
  parallel = FALSE)
```


```{r, eval = T, echo=TRUE}
sum(round(solution1$aggr_strata$SOLUZ))
expected_CV(solution1$aggr_strata)
```

## Optimization without taking into account the models

We now evaluate the expected precision on the true target variables:

```{r, eval = T, echo=TRUE}
newstrata <- updateStrata(swissstrata1,solution1)
framenew1 <- updateFrame(swissframe,newstrata)
framenew1 <- framenew1[order(framenew1$ID),]
swissmunicipalities <- swissmunicipalities[order(swissmunicipalities$id),]

frame_pop <- framenew1
frame_pop$Y1 <- swissmunicipalities$Airbat
frame_pop$Y2 <- swissmunicipalities$Pop020
frame_pop$Y3 <- swissmunicipalities$Pop2040
eval1 <- evalSolution(frame_pop, 
                     solution1$aggr_strata, 
                     nsampl=100, 
                     writeFiles=FALSE,
                     progress=FALSE) 
eval1$coeff_var
```

## Optimization taking into account the models

Let us now build the atomic strata taking into account the fitted models:


```{r, eval = T, echo=TRUE}
swissstrata2 <- buildStrataDF(swissframe, model = model, progress = FALSE)
```

## Optimization taking into account the models

and then proceed with the optimization:

```{r, eval = T, echo=TRUE}
set.seed(123)
solution2 <- optimizeStrata(
  errors = swisserrors, 
  strata = swissstrata2,
  iter = 50,
  pops = 10,
  writeFiles = FALSE,
  showPlot = FALSE,
  parallel = FALSE)

```

```{r, eval = T, echo=TRUE}
sum(round(solution2$aggr_strata$SOLUZ))
expected_CV(solution2$aggr_strata)
```

## Optimization taking into account the models

Again, let us evaluate the expected precision on the true target variables:

```{r, eval = T, echo=TRUE}
newstrata <- updateStrata(swissstrata2,solution2)
framenew2 <- updateFrame(swissframe,newstrata)
framenew2 <- framenew2[order(framenew2$ID),]
swissmunicipalities <- swissmunicipalities[order(swissmunicipalities$id),]
# evaluation of precision on real target variables

frame_pop <- framenew2
frame_pop$Y1 <- swissmunicipalities$Airbat
frame_pop$Y2 <- swissmunicipalities$Pop020
frame_pop$Y3 <- swissmunicipalities$Pop2040
eval2 <- evalSolution(frame_pop, 
                     solution2$aggr_strata, 
                     nsampl=100, 
                     writeFiles=FALSE,
                     progress=FALSE) 
eval2$coeff_var
swisserrors

```

